---
title: "Cluster Model: Downsampling with hvac_dhw_var"
author: "Noah Klammer"
date: "6/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Clear global env and report

```{r include=FALSE}
rm(list = ls())
gc()
```

# Intro

This is my thrid attempt at clustering of zones using the ideal air loads output variable. First I tried 3-space, then I  tried 364-space clustering. Confident in the speed of computation, I'm going to try all hours in the simulation year, 8760-space clustering.

## Import Rda data

```{r import, message=FALSE, warning=FALSE}
load(file = "hvac_dhw_var.Rda")
```

### Set data

```{r}
# CHANGE DATA HERE
df <- hvac_dhw_var
```

## Downsample day, week, month

Here we can choose to downsample our temporal data from hour to `day`, `week`, or `month`.

```{r}
downsample <- TRUE
sample_interval <- "month" # choices: "day", "week", "month

if (downsample == TRUE) {
  if (sample_interval == "day") {
    # downsample to day
    day_idx <- data.frame(day_idx = rep(1:365, each = 24))
    rownames(day_idx) <- rownames(df)
    df <- cbind(day_idx,df) %>%
      group_by(day_idx) %>%
      summarize(across(-c(day,hour,month),sum)) %>%
      select(-c(day_idx))
    hvac_dhw_var_day <- df
    save(hvac_dhw_var_day, file = "hvac_dhw_var_day.Rda")
  }
  else if (sample_interval == "week") {
  # downsample to week
  week_idx <- data.frame(week_idx = rep(1:52, each = 7*24, length.out = 8760))
  rownames(week_idx) <- rownames(df)
  df <- cbind(week_idx,df) %>%
    group_by(week_idx) %>%
    summarize(across(-c(day,hour,month),sum)) %>%
    select(-c(week_idx))
  hvac_dhw_var_week <- df
  save(hvac_dhw_var_week, file = "hvac_dhw_var_week.Rda")
  }
  else {
  # downsample to month
  df <- df %>%
    group_by(month) %>%
    summarize(across(-c(day,hour),sum)) %>%
    select(-c(month))
  hvac_dhw_var_month <- df
  save(hvac_dhw_var_month, file = "hvac_dhw_var_month.Rda")
  }
}
  
```

# Feature normalization

Normalize values by floor area of the respective zone.

## Area to Zone mapping

Load in an external data set of rows with two attributes: zone name and floor area in m^2. Since the zone-area dataset and the Ideal Air Loads dataset may have different order, create a simple indexing function.

### Normalize Loads by Floor Area

```{r message=FALSE}

### Drop Date/Time and save for later
if ("hour" %in% colnames(df)) {
  time_cols <- select(df, c(hour,day,month))
  rownames(time_cols) <- rownames(df) # need to keep row labels
  df <- select(df, -c(hour,day,month)) 
} else if ("day" %in% colnames(df)) {
  time_cols <- select(df, c(day,month))
  rownames(time_cols) <- rownames(df) # need to keep row labels
  df <- select(df, -c(day,month)) 
} else if ("month" %in% colnames(df)) {
  time_cols <- select(df, month)
  rownames(time_cols) <- rownames(df)
  df <- select(df, -month)
} else {}

### Load Area Map
area_map <- read_csv("../ZoneFloorArea-Map.csv")

idx_f <- function(string) { # takes zone string and maps to zone index in area_map
  which(string==area_map$`Zone List`) # returns num vec
  }

area_idx <- sapply(colnames(df),idx_f) # maps df idx to area idx
area_vec <- area_map$`Space Area [m2]`[area_idx]

# apply normalization vector across columns
df <- sweep(df, 2, area_vec, FUN = "/") 

# change units
# **[J]** => [J/m^2]
rownames(df) <- str_replace(rownames(df),"(?<=\\[)J{1}","J/m^2")

### Add Date/Time back in
if (exists("time_cols")) {
df <- cbind(time_cols,df)
}
```


# Clustering

Traditionally, we would remove columns with zero variance as they are unhelpful in the sense of regression. However, in clustering we may want to leave them in.



## Transpose and cluster

Let's introduce the `apcluster` package which is an implementation of Frey and Dueck's popular Affinity Propagation method for passing messages between pairs of data. I would make sure to reference the [math paper](https://doi.org/10.1080/19401493.2017.1410572), the [R package](https://doi.org/10.1093/bioinformatics/btr406), and the [original method's](https://doi.org/10.1126/science.1136800) publication.


```{r turnkey cluster process}
library(apcluster)

# drop time date cols in
# preparation for clustering
if ("minute" %in% colnames(df)) {
  df <- subset(df, select = -c(day, hour, month, minute))
} else if ("day" %in% colnames(df)) {
  df <- subset(df, select = -c(day, hour, month))
} else {}

tdf <- as.data.frame(t(df))

APR <- apcluster(negDistMat(r=2), tdf, details = TRUE) # returns a APResult

area_map <- read_csv("../ZoneFloorArea-Map.csv")

idx_f <- function(string) { # takes zone string and maps to zone index in area_map
  which(string==area_map$`Zone List`) # returns num vec
  }

# this creates the scalar vector
# but does not apply the scalar on the df yet
scalars <- vector(mode = "numeric")
for (i in 1:length(APR@clusters)) { # iterate through each cluster
  # get list of members of cluster i # char vec
  member_zone_names <- names(unlist(APR@clusters[i])) # inclusive of exemplar
  # map strings to m2 values # num vec
  area_idx <- sapply(member_zone_names, idx_f) # maps cluster idx to area idx
  # get floor area num values
  member_area_num <- area_map$`Space Area [m2]`[area_idx] # num vec
  # create scalar and append to num vec in order of clusters i
  # sum area numbers
  scalars <- append(scalars, sum(member_area_num)) # scaling factor is sum of areas
}

# reduce dimensionality of 'df' using clusters
red_df <- df[APR@exemplars] # returns 8760 rows with reduced columns

# apply scalar vec from above to red_df
# is the scalar of the right length?
ncol(red_df) == length(scalars)

# apply area scalars to reduced df
red_df <- sweep(red_df, 2, scalars, FUN = "*")

# change units
rownames(red_df) <- str_replace(rownames(red_df),"J/m\\^2","J")

### Add Date/Time back in
if (exists("time_cols")) {
red_df <- cbind(time_cols,red_df)
}

### Save out
if (sample_interval=="day") {
  hvac_dhw_var_day_red <- red_df
  save(hvac_dhw_var_day_red, file = "hvac_dhw_var_day_red.Rda")
  print("The file 'hvac_dhw_var_day_red.Rda' was saved.")
} else if (sample_interval=="week") {
  hvac_dhw_var_week_red <- red_df
  save(hvac_dhw_var_week_red, file = "hvac_dhw_var_week_red.Rda")
  print("The file 'hvac_dhw_var_week_red.Rda' was saved.")
} else if (sample_interval=="month") {
  hvac_dhw_var_month_red <- red_df
  save(hvac_dhw_var_month_red, file = "hvac_dhw_var_month_red.Rda")
  print("The file 'hvac_dhw_var_month_red.Rda' was saved.")
} else {print("No file was saved.")}

```

# End

<br><br><br>