---
title: "Final Results B case"
author: "Noah Klammer"
date: "6/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Clear global env and report

```{r include=FALSE}
rm(list = ls())
gc()
```

# Intro

In this case B, I am comparing a reduced model with clustering basis in `ilas_dhw_novar`, with a full model `hvac_dhw_novar`.

## Import Rda data

```{r import, message=FALSE, warning=FALSE}
load(file = "hvac_dhw_novar.Rda")
load(file = "compare_B.Rda")

```

### Set data

```{r}
# CHANGE DATA HERE
df <- hvac_dhw_novar
df_red <- compare_B
```

# Body

### Sum across rows

```{r}
# maybe summarize verb across rows?

df <- df %>%
  replace(is.na(.), 0) %>%
  mutate(rowsum = rowSums(.)) %>%
  select(rowsum)

df_red <- df_red %>%
  replace(is.na(.), 0) %>%
  mutate(rowsum = rowSums(.)) %>%
  select(rowsum)

hour_index <- seq(8760)
red_rowsum <- df_red$rowsum
df_com <- cbind(hour_index,df,red_rowsum) # comparison df

```


### Definitions of CV-RMSE and NMBE

I am coding working backwards from results. For equations of CV-RMSE and NMBE see, Results chapter.

```{r}
cvrmse <- function(y_i, y_hat_i) {
  sqrt(sum(((y_i - y_hat_i)^2))/(length(y_i)-1))/mean(y_i)
}
nmbe <- function(y_i, y_hat_i) {
  sum(y_i-y_hat_i)/((length(y_i)-1)*mean(y_i))
}
```

# Do stat calcs

```{r}
cvrmse(df_com$rowsum, df_com$red_rowsum)
nmbe(df_com$rowsum, df_com$red_rowsum)
sum(df_com$red_rowsum)/sum(df_com$rowsum)
```

## Plot the residuals

```{r}
plot(df_com$rowsum, df_com$red_rowsum, pch = 20, ylab="Energy use, reduced model (J)", xlab="Energy use, full model (J)", main="Reduced Model versus Full Model")
abline(0,1, col = "red")

```


```{r}
residuals = df_com$rowsum - df_com$red_rowsum

hour_index <- seq(8760)
#plot(hour_index, residuals, ylab="Residuals", xlab="Year hours", main="Reduced Model versus Full Model") 
#abline(0, 0, col="red")

qqnorm(residuals, frame = FALSE, main = "Normal Q-Q Plot of Residuals")
qqline(residuals, col = "red")


```


# Downsample day, week, month

```{r}
# downsample to day
day_idx <- rep(1:365, each = 24)

cbind(day_idx,df) %>%
  group_by(day_idx) %>%
  summarize(across(-c(day,hour,month),sum))

# downsample to week
week_idx <- rep(1:52, each = 7*24, length.out = 8760)
cbind(week_idx,df) %>%
  group_by(week_idx) %>%
  summarize(across(-c(day,hour,month),sum))

# downsample to month
df %>%
  group_by(month) %>%
  summarize(across(-c(day,hour),sum))

  
```

# End

<br><br><br>